# REPRODUCIBLE RESEARCH COURSERA ONLINE COURSE
## WEEK 2 PEER-REVIEW PROJECT

## Reading the data
```{r}
act <- read.csv("activity.csv")
```

## Mean steps taken per day
```{r}
meansteps <- tapply(act$steps, act$date, mean, na.rm=TRUE)
meansteps
```

## Number of steps taken per day

```{r}
stepsperday <- tapply(act$steps, act$date, sum, na.rm=TRUE)
stepsperday
```

## Histogram steps taken each day

```{r}
hist(stepsperday, xlab = "Steps per day", main = "Histogram Steps Taken Per Day")
```

## Mean and Median Steps taken per day

```{r}
meanstepsperday <- mean(stepsperday)
medianstepsperday <- median(stepsperday)
```

The mean of the total number od steps taken per day is 9354.23,
The median of the total number of steps taken per day is 10395.

## Time-series Plot

# Preliminary wrangling

```{r}
library(lubridate)
act$date <- ymd(act$date)
```

```{r}
s <- split(act, act$date)
meanofstepsperday <- sapply(s, function(x){mean(x[,1],na.rm = TRUE)})

##Function to help rearrange dates
formatdate <- function(x){
        nx <- nrow(x)
        newdate <- numeric(nx)
        for (i in 1:nx) {
                if(nchar(x[i,3])==1){x[i,3] <- paste0(000,x[i,3])}
                if(nchar(x[i,3])==2){x[i,3] <-paste0(00,x[i,3])}
                if(nchar(x[i,3])==3){x[i,3] <-paste0(0,x[i,3])}
                
        x$newdate[i] <- paste(x[i,2],x[i,3])
        }
        x
}

newact <- formatdate(act)

newdf <- data.frame(averagesteps=meanofstepsperday, date=unique(newact$date))

with(newdf, plot(date, averagesteps, type="l", col="magenta"))
```

## Missing values

# The total number of missing values.

```{r}
sum(is.na(act[,1:3]))
```

# Imputting Missing Values using the means of intervals

```{r}
act1 <- act
act1$interval <- as.factor(act1$interval)
```

# A function to implement the strategy

```{r}
imputna <- function(x){
        nx <- nrow(x)
        mean_by_interval <- tapply(act1[,1], act1[,3], mean, na.rm= T)
        m <- as.vector(mean_by_interval)
        x[1:288,1] <- m[1:288]; x[2017:2304,1] <- m[1:288]
        x[8929:9216,1] <- m[1:288]; x[9793:10080,1] <- m[1:288]
        x[11233:11520,1] <- m[1:288]; x[11521:11808,1] <- m[1:288]
        x[12673:12960,1] <- m[1:288]; x[17281:17568,1] <- m[1:288]
        
        x
}
```

# New data set with missing values added

```{r}
act2 <- imputna(act1)
```

## Total number of steps taken each day and the histogram

```{r}
sumt <- tapply(act2$steps, act2$date, sum)
hist(sumt)
```

## Median and Mean total number of steps taken each day

```{r}
mediant <- tapply(act2$steps, act2$date, median)
meant <- tapply(act2$steps, act2$date, mean)
```


## Weekdays vs Weekends

```{r}
act2$week <- factor(weekdays(act2$date))
```

```{r, results= "hide"}
library(tidyverse)
```

```{r}
act2$weekdayorend <- recode(act2$week, "Monday"="weekday", "Tuesday"="weekday",
"Wednesday"="weekday","Thursday"="weekday","Friday"="weekend",
"Saturday"="weekend","Sunday"="weekend")

##Panel plot
ggplot(act2, aes(interval, steps))+ geom_line()+ facet_wrap(~weekdayorend)
```

